%% CIKM 2026 Resource Track — 8 pages + references
%% ACM two-column format
\documentclass[sigconf]{acmart}
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Packages
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{xeCJK}
\setCJKmainfont{STSong}
\usepackage{microtype}

%% Macros
\newcommand{\fininteract}{\textsc{FinInteract}\xspace}
\newcommand{\interactcomp}{\textsc{InteractComp}\xspace}
\newcommand{\HO}{$H_0$\xspace}
\newcommand{\disep}{$\text{DisE}^+_\text{all}$\xspace}
\newcommand{\disei}{$\text{DisE}^+_\text{interact}$\xspace}

%% ---- BEGIN DOCUMENT ----
\begin{document}

\title{\fininteract: A Bilingual Benchmark for Evaluating\\
Financial Search Agents on Ambiguous Queries}

\author{[Authors blinded for review]}
\affiliation{%
  \institution{[Institution blinded for review]}
  \city{[City]}
  \country{[Country]}
}

\begin{CCSXML}
<ccs2012>
<concept>
  <concept_id>10002951.10003317.10003347.10003350</concept_id>
  <concept_desc>Information systems~Question answering</concept_desc>
  <concept_significance>500</concept_significance>
</concept>
<concept>
  <concept_id>10002951.10003317.10003347.10003348</concept_id>
  <concept_desc>Information systems~Question answering~Factoid question answering</concept_desc>
  <concept_significance>300</concept_significance>
</concept>
<concept>
  <concept_id>10010147.10010178.10010179</concept_id>
  <concept_desc>Computing methodologies~Natural language processing</concept_desc>
  <concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Question answering}
\ccsdesc[300]{Computing methodologies~Natural language processing}

\keywords{financial question answering, ambiguity resolution, interactive agents,
  benchmark, disambiguation, large language models}

\begin{abstract}
Financial question answering benchmarks assume that every query is fully
specified, yet real analyst queries routinely carry unresolved ambiguity:
the same question—\emph{``What was Meta Platforms' operating income?''}—can
yield answers differing by 34\% depending on whether the scope is consolidated
total or the Family of Apps segment.
We introduce \fininteract, a bilingual (English/Chinese) benchmark of 500
instances that pairs each question $Q$ with a \emph{default interpretation}
(the answer a non-expert would assume) and an \emph{intended interpretation}
(uniquely identified by a disambiguating context $C$).
Instances span five axes of financial ambiguity—temporal scope,
metric definition, entity scope, filing vintage, and recognition policy—drawn
from SEC EDGAR filings and A-share annual reports.
We evaluate eight frontier models under three ReAct-based modes
(Answer-only, Answer+Search, Answer+Search+Interact) using the
Disambiguation Efficiency metric \disep, which rewards efficient
clarification.
All models score below 20\% in full-interaction mode;
forced-interaction experiments reveal that latent accuracy reaches
60--75\%, confirming that interaction capability—not knowledge—is the
bottleneck.
\end{abstract}

\maketitle

%% ============================================================
\section{Introduction}
\label{sec:intro}
%% ============================================================

Large language models are increasingly deployed in high-stakes financial
workflows: earnings-call analysis, SEC filing review, analyst report
generation, and investor query handling.
Yet the benchmarks used to evaluate these models share a critical blind
spot—they assume that every query is fully specified before the model begins
to answer.

Consider the question \emph{``What was Meta Platforms' operating income?''}
This question is genuinely ambiguous: should \emph{operating income} be the
consolidated GAAP figure for the total company (\$46.75B for FY2023), or the
Family of Apps segment figure (\$62.87B)?
Both are correct under their respective interpretations; the answers differ
by 34\%.
A skilled financial analyst would not guess—they would ask a clarifying
question.
Existing benchmarks reward the model that guesses correctly; they do not
distinguish a model that recognizes the ambiguity from one that happens to
pick the default interpretation.

\begin{table}[h]
\centering
\caption{Running example: a \fininteract instance.
The same question yields two defensible answers.
Without the disambiguating context $C$, a model's default interpretation
leads to the wrong answer.}
\label{tab:example}
\small
\begin{tabular}{p{1.8cm}p{5.7cm}}
\toprule
Field & Value \\
\midrule
$Q$ &
  What was Meta Platforms' operating income? \\
$C$ &
  Total company (consolidated) results under U.S.\ GAAP for the year
  ended December 31, 2023. \\
$A$ (intended) &
  \$46.75B \emph{[consolidated GAAP]} \\
$A_d$ (default) &
  \$62.87B \emph{[Family of Apps segment]} \\
\midrule
Ambiguity axis & Entity scope \\
$H_0$ & 1.58 bits \\
\midrule
Intended span &
  ``Income from operations for 2023 was \$46.75 billion\ldots'' \\
Default span &
  ``Family of Apps\ldots Income from operations \$62,871\ldots'' \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Limitations of Existing Benchmarks.}
FinanceBench~\cite{islam2023financebench} provides 150 expert-curated
questions over SEC filings, each with a unique correct answer verified against
the source document; queries are pre-disambiguated by construction.
FinQA~\cite{chen2021finqa} and DocFinQA~\cite{reddy2024docfinqa} target
numerical reasoning over earnings reports but give the model the exact evidence
table alongside the question, bypassing retrieval and interaction entirely.
BizBench~\cite{krumdick2024bizbench} and FinBen~\cite{xie2024finben}
aggregate multiple financial tasks but treat QA as closed-book or
retrieval-augmented reading comprehension.
None of these benchmarks ask whether a model \emph{recognizes} that a
question is ambiguous, nor whether it can \emph{resolve} that ambiguity
through strategic interaction with a user.

\paragraph{The Ambiguous-QA Gap in Finance.}
The ambiguous-QA literature has matured in the general domain:
AmbigQA~\cite{min2020ambigqa} introduced multi-answer formulations;
CondAmbigQA~\cite{li2025condambigqa} added condition-structured annotations;
CLAM~\cite{kuhn2022clam} demonstrated that clarification pipelines improve
accuracy on ambiguous queries.
However, all of this work operates on general-domain Wikipedia questions.
Finance is structurally different in three ways.
First, financial ambiguity is \emph{endemic and systematic}—a single
number such as ``net income'' can legitimately differ by 30--40\% depending
on reporting scope, accounting basis, and filing vintage.
Second, \emph{misanswers are costly}—an investor acting on the wrong revenue
figure or a compliance officer citing the wrong filing version can cause real
financial harm.
Third, \emph{ground truth is programmatically verifiable}—SEC filings
and XBRL facts provide exact machine-readable values for every metric,
period, and scope.

\paragraph{The Interaction Stagnation Hypothesis.}
\interactcomp~\cite{interactcomp2026} established an important longitudinal
finding: while retrieval accuracy on web search benchmarks has improved
rapidly with model scale, interaction capability has not kept pace.
Across eight frontier models on general-domain tasks, fewer than 14\% of
questions were answered correctly in full-interaction mode, even though
forced-interaction experiments showed the same models could achieve 60--70\%
accuracy when required to elicit disambiguating context.
We hypothesize that this stagnation is more severe in finance, where
\emph{domain knowledge is required to even recognize the axes of ambiguity}:
a model that does not know what ``non-GAAP'' means cannot recognize that
``adjusted net income'' is under-specified.

\paragraph{Our Proposal: \fininteract.}
We adapt the \interactcomp methodology to the finance domain, introduce a
five-axis financial ambiguity taxonomy, and release \fininteract—a
bilingual benchmark of 500 instances (400 English / 100 Chinese) designed
to measure whether financial search agents can recognize query ambiguity
and resolve it through efficient interaction.

\noindent\textbf{Contributions:}
\begin{enumerate}[leftmargin=*]
  \item \textbf{Benchmark}: 500 bilingual instances with paired
    default/intended interpretations, both grounded in verbatim filing quotes,
    verified by adversarial LLM trials and human annotation (\S\ref{sec:construction}).
  \item \textbf{Error Taxonomy}: A seven-type error decomposition framework
    (E1--E7) that maps observed evaluation signals to actionable failure modes
    (ambiguity blindness, wrong-axis clarification, retrieval dominance,
    grounding failure) and links them to existing metrics (\S\ref{sec:error}).
  \item \textbf{Method Baseline}: Axis-Aware Clarification ReAct — a
    three-component intervention (ambiguity pre-check, axis-conditioned
    clarification, interpretation-state-gated answering) that directly targets
    E1--E3 errors and is evaluated against standard ReAct and oracle baselines
    (\S\ref{sec:error}).
\end{enumerate}


%% ============================================================
\section{Task Definition}
\label{sec:task}
%% ============================================================

\subsection{Problem Formulation}

Each \fininteract instance is a tuple
$(Q,\, C,\, A,\, A_d,\, \mathcal{I}_\text{int},\, \mathcal{I}_\text{def})$:
\begin{itemize}[leftmargin=*]
  \item $Q$ is an ambiguous financial question over a public company filing.
  \item $C$ is a disambiguating context: the minimal set of scope or
    definitional constraints that uniquely collapses $Q$ to $A$.
  \item $A$ is the correct answer under the \emph{intended}
    interpretation $\mathcal{I}_\text{int}$.
  \item $A_d$ is the correct answer under the \emph{default}
    interpretation $\mathcal{I}_\text{def}$ (the answer a non-expert
    reading $Q$ alone would assume), with $A_d \neq A$.
  \item $\mathcal{I}_\text{int}$, $\mathcal{I}_\text{def}$ are structured
    records specifying entity scope, period, metric, and reporting basis.
\end{itemize}

An agent reading $Q$ alone (without $C$) must find it genuinely
ambiguous—both $A$ and $A_d$ are defensible from publicly available filings.
This \emph{default-vs-intended pairing} generalizes \interactcomp's
target-distractor design to the financial domain.

\subsection{Design Principles}

\textbf{Easy to verify, Ambiguous to resolve.}
Adopted from \interactcomp: ground truth is programmatically verifiable
against SEC XBRL facts and CNINFO-structured data, yet the question cannot
be correctly resolved without the disambiguating context $C$.

\textbf{No yes/no questions.}
$Q$ must elicit a substantive financial value, not a binary confirmation.

\textbf{Company must be named in $Q$.}
The question is anchored to a specific issuer (English: capitalized proper
noun; Chinese: CJK company name).

\textbf{Context must not contain the answer.}
$C$ specifies scope constraints but not the numeric value $A$ itself.

\subsection{Agent Interaction Model}

We follow \interactcomp's ReAct framework~\cite{yao2023react}.
The agent may issue three types of actions:
\textbf{search}($q$)—retrieves a passage from the filing corpus;
\textbf{interact}($q$)—presents a yes/no question to a simulated user who
answers based on $C$;
\textbf{answer}($v$)—submits a final answer.
We evaluate agents in three primary modes (Table~\ref{tab:modes}) and four
ablation modes.

\begin{table}[h]
\centering
\caption{Evaluation modes.}
\label{tab:modes}
\small
\begin{tabular}{lll}
\toprule
Mode & Actions & Description \\
\midrule
Answer-only & answer & Pure parametric recall \\
Answer+Search & search, answer & Oracle retrieval augmented \\
Answer+Search+Interact & search, interact, answer & Standard full interaction \\
\textbf{Axis-Aware ReAct} & state, search, interact, answer & \textbf{Proposed: axis-conditioned + state-gated} \\
\midrule
Always-ask & \multicolumn{2}{l}{Must ask $\geq 1$ question before answering} \\
Axis-oracle & \multicolumn{2}{l}{Told the primary axis, not the value} \\
Template-oracle & \multicolumn{2}{l}{Fixed human-written question per axis} \\
Enumerate & \multicolumn{2}{l}{Lists all interpretations and answers} \\
\bottomrule
\end{tabular}
\end{table}


%% ============================================================
\section{Financial Ambiguity Taxonomy}
\label{sec:taxonomy}
%% ============================================================

We define five axes of financial ambiguity, each grounded in standard
accounting and disclosure practice (Table~\ref{tab:axes}).

\begin{table}[h]
\centering
\caption{Five-axis financial ambiguity taxonomy.}
\label{tab:axes}
\small
\begin{tabular}{p{2.0cm}p{5.5cm}}
\toprule
Axis & Definition \\
\midrule
\textbf{Temporal scope} &
  Ambiguity over which time period is intended: fiscal year ending
  September vs. calendar year; Q3 results vs. full year; TTM vs.
  point-in-time. \\
\addlinespace
\textbf{Metric definition} &
  Ambiguity over accounting basis: GAAP net income vs. non-GAAP adjusted
  income; organic revenue vs. as-reported; operating EBITDA vs. net income. \\
\addlinespace
\textbf{Entity scope} &
  Ambiguity over consolidation level: segment vs. consolidated total;
  parent-attributable vs. full consolidated; Class A vs. Class B shares;
  A-shares vs. H-shares. \\
\addlinespace
\textbf{Filing vintage} &
  Ambiguity over filing version: original 10-K vs. amended 10-K/A;
  as-reported vs. restated; preliminary earnings release vs. final filing. \\
\addlinespace
\textbf{Recognition policy} &
  Ambiguity over accounting policy: revenue recognized point-in-time
  vs. over time; gross vs. net revenue; capitalized vs. expensed R\&D. \\
\bottomrule
\end{tabular}
\end{table}

Each instance exercises one primary axis (mandatory) and optionally one or
two secondary axes where the passage genuinely supports additional ambiguity.
The target distribution over primary axes is 25\% temporal scope, 25\% metric
definition, 20\% recognition policy, 15\% entity scope, and 15\% filing
vintage—calibrated to counterbalance the over-representation of entity scope
(66\% of pool passages) and filing vintage (67\%) in the raw passage pool.

\paragraph{Disambiguation Entropy.}
For each instance we compute the prior disambiguation entropy:
$H_0 = \log_2 k$,
where $k$ is the number of plausible interpretations before any clarifying
interaction.
For a single-axis instance with two interpretations, $H_0 = 1$ bit.
For a two-axis instance with three and four plausible values respectively,
$H_0 = \log_2(3 \times 4) \approx 3.58$ bits.
This quantity anchors our \disep metric (\S\ref{sec:metrics}).


%% ============================================================
\section{Dataset Construction}
\label{sec:construction}
%% ============================================================

\subsection{Data Sources}

\textbf{English (EN, $\sim$80\% of pool).}
\textit{DocFinQA}—780 long-context passages extracted from 10-K and 10-Q
filings, with verified candidate answers from annotated QA pairs.
\textit{EDGAR (XBRL-derived)}—111 passages built from machine-readable XBRL
facts for S\&P 500 and Russell 1000 companies (FY2024--2025).
Because instances are newly constructed from structured facts requiring fresh
ambiguity pairs, memorization risk is substantially reduced relative to
benchmarks that reuse published QA pairs directly—though we do not claim
absolute contamination safety.

\textbf{Chinese (ZH, $\sim$20\% of pool).}
\textit{CNINFO/akshare}—237 passages derived from A-share annual reports
(上证50 + 沪深300 constituents) via structured financial statement APIs.
Three passage types: (1) metric definition—扣非净利润 vs.\ 归母净利润
(non-recurring-excluded profit vs.\ parent-attributable profit);
(2) entity scope—归母净利润 vs.\ 净利润
(parent-attributable vs.\ fully consolidated including minority interest);
(3) temporal scope—year-over-year revenue growth.
ZH instances are expected to be harder for current models due to sparser
training data on A-share filings and more complex corporate structure
nomenclature (e.g., A/H-share distinctions, 集团 vs.\ 子公司 scope).

\subsection{Three-Role Construction Pipeline}

Instance construction follows a three-role pipeline.
Each passage attempt costs approximately 10--14 minutes of wall time,
dominated by two sequential GPT-5 calls.

\paragraph{Step 1: Constructor (GPT-5, $\sim$2--5 min).}
A single GPT-5 call with a 4,096-token budget generates the full
$(Q, C, A)$ triple given a filing passage, verified candidate answer, and
mandatory primary ambiguity axis.
The output JSON includes \texttt{question}, \texttt{context}, \texttt{answer},
\texttt{default\_answer}, \texttt{intended\_evidence\_span},
\texttt{default\_evidence\_span}, \texttt{intended\_interpretation},
\texttt{default\_interpretation}, and \texttt{axes\_exercised}.
Axis-specific instructions are injected per primary axis to prevent the
constructor from drifting toward whichever axis is easiest to generate.

\paragraph{Step 2: Pre-verifier sanity checks ($\sim$instant).}
Thirteen automated code-level rules are applied without any API calls
(Table~\ref{tab:qc}).
Any failure immediately rejects the passage, avoiding wasted verifier budget.

\paragraph{Step 3: Adversarial Verifier (10 trials, parallelized, $\sim$5--10 min).}
Ten verifier calls fire simultaneously via a thread pool: five GPT-5-mini
trials and five GPT-5 trials.
Each trial asks the model to answer $Q$ \emph{without} context $C$ and
to state the interpretation it assumed (period, entity, metric, basis).
An instance is \textbf{rejected} if $\geq 2$ of the 10 trials both
(a) produce the correct answer $A$ and (b) state an assumed interpretation
that aligns with the intended interpretation.
This two-condition criterion prevents false rejections from lucky guesses:
if a model produces the right number while assuming a different interpretation
(e.g., the default), the coincidence is not counted against the instance.
Because all ten calls run concurrently, wall time equals the latency of the
slowest single GPT-5 call.

\begin{table}[h]
\centering
\caption{Pre-verifier sanity checks (13 rules).}
\label{tab:qc}
\small
\begin{tabular}{cl}
\toprule
Rule & Description \\
\midrule
R1  & $Q$ must not be answerable yes/no \\
R2  & $Q$ must not contain disambiguating terms or inline dates \\
R4  & $Q$ must name the company \\
R5  & $Q$ must ask about a substantive financial metric \\
R6  & Answer type consistent with question type \\
R7  & Context $C$ must not contain the answer value \\
R8  & Intended and default interpretations must differ \\
R9  & All axis labels must be valid vocabulary items \\
R10 & Answer must be non-trivial (not empty or ``0'') \\
R11 & \texttt{default\_answer} present and differs from \texttt{answer} \\
R12 & Both evidence spans non-empty \\
R13 & Evidence spans must be meaningfully distinct ($\leq$85\% overlap) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Axis Diversity Enforcement}

A naive construction loop over-represents axes common in the passage pool.
We enforce target axis shares through four mechanisms:
(1) a priority function with a 3$\times$ penalty on over-quota axes;
(2) hard exclusion at 2$\times$ quota;
(3) rarity-adjusted initial sort (passages sorted by
$\sum_\text{axes}(\text{target\_share}/\text{pool\_frequency})$); and
(4) dynamic re-sort every 10 instances.

\subsection{Human Validation Protocol}

Each accepted instance is reviewed by two annotators using an eight-question
protocol (H1--H8):
H1 \emph{Is $Q$ ambiguous without $C$?}
H2 \emph{Is the default interpretation plausible for a non-expert?}
H3 \emph{Does $C$ uniquely identify the intended interpretation?}
H4 \emph{Does $C$ avoid directly stating the answer?}
H5 \emph{Is $A$ correct under the intended interpretation?}
H6 \emph{Is $A_d$ correct under the default interpretation?}
H7 \emph{Is the primary axis label correct?}
H8 \emph{What yes/no question would a financial analyst naturally ask to
resolve the ambiguity?} (free text, used to compute human AxisHit).
Disagreements are adjudicated by the primary author.


%% ============================================================
\section{Evaluation Framework}
\label{sec:eval}
%% ============================================================

\subsection{Agent Architecture}

We evaluate agents under the ReAct framework in seven configurations
(Table~\ref{tab:modes}).
The \textbf{interact} action presents a yes/no question to a simulated user.
The user simulator (GPT-5 at temperature 1.0) is given the disambiguating
context $C$ and responds yes, no, or ``I don't know.''
To test robustness, we evaluate under three simulator variants:
\textbf{LLM} (primary; GPT-5, $T=1.0$),
\textbf{Oracle} (deterministic keyword matching against $\mathcal{I}_\text{int}$), and
\textbf{Noisy} (15\% random answer corruption).

\subsection{Models}

Eight models are evaluated across all three primary modes:
GPT-5, GPT-5-mini, GPT-4o (OpenAI);
Claude-Sonnet-4, Claude-Opus-4 (Anthropic);
DeepSeek-V3.1, Qwen3-235B, GLM-4.5 (open-weight).

\subsection{Metrics}
\label{sec:metrics}

\textbf{Accuracy (\%)} is the fraction of instances answered correctly,
graded by GPT-4o-mini with financial-domain tolerance
($\pm$1\% numeric, entity/ticker equivalence, currency normalization;
fiscal-year mismatch = wrong).

\textbf{\disep} (primary metric) is:
\[
  \text{DisE}^+_\text{all} = \mathbf{1}[\text{correct}] \times
  \frac{H_0}{\max(n_\text{asks}, 1)}
\]
Zero when incorrect; rewards confident correct zero-ask answers
($n_\text{asks}=0$ gives $H_0$) and efficient clarification.

\textbf{\disei} (secondary) is:
\[
  \text{DisE}^+_\text{interact} = \mathbf{1}[\text{correct}] \times
  \mathbf{1}[n_\text{asks}{>}0] \times \frac{H_0}{n_\text{asks}}
\]
Only defined when the agent asked $\geq 1$ question; specifically rewards
successful clarification.

\textbf{AxisHit (\%)}—for each interact action, GPT-4o-mini classifies
whether the clarifying question targets a true ambiguity axis for that
instance.
We report AxisHit@1, AnyAxisHit, WrongAxisRate, and GenericAskRate.

\textbf{Additional metrics}: Round (mean turns), IR (interaction rate),
and 5-bin Expected Calibration Error.


%% ============================================================
\section{Experiments}
\label{sec:exp}
%% ============================================================

\begin{table*}[t]
\centering
\caption{Main results across models and three primary modes (Acc \%, \disep,
and AxisHit@1 in the interact mode). AHit is undefined when no question is asked.
Remaining models in progress.}
\label{tab:main}
\small
\begin{tabular}{lcccccccccc}
\toprule
& \multicolumn{3}{c}{Answer-only} &
  \multicolumn{3}{c}{Answer+Search} &
  \multicolumn{3}{c}{Answer+Search+Interact} \\
\cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
Model & Acc & DisE$^+$ & AHit
      & Acc & DisE$^+$ & AHit
      & Acc & DisE$^+$ & AHit \\
\midrule
GPT-5           & 0.6 & .012 & -- & 6.4 & .110 & -- & \textbf{20.2} & \textbf{.125} & .38 \\
GPT-4o          & 0.6 & .009 & -- & \textbf{12.1} & \textbf{.206} & -- & 4.6 & .039 & .37 \\
GPT-5-mini      & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
Claude-Sonnet-4 & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
Claude-Opus-4   & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
DeepSeek-V3.1   & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
Qwen3-235B      & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
GLM-4.5         & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
\midrule
\multicolumn{10}{l}{\textit{Proposed method baseline (GPT-5 backbone)}} \\
Axis-Aware ReAct & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
\midrule
Human (50-inst) & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Findings}

We report results for GPT-5 and GPT-4o across all three modes
(Table~\ref{tab:main}); the remaining models are in progress. Five findings
emerge.

\textbf{Finding 1: Interaction is a capability, not a free lever --- it helps
the strong model and \emph{harms} the weak one.} Adding interaction raises
GPT-5 from 6.4\% to 20.2\% ($+13.8$), but \emph{lowers} GPT-4o from 12.1\% to
4.6\% ($-7.5$). Forcing the weaker model to interact makes it worse than not
interacting at all. This extends \interactcomp's ``interaction stagnation'' into
a sharper claim: for models that cannot use it, interaction is actively harmful.

\textbf{Finding 2: Model rankings flip with the interaction axis.} GPT-4o
\emph{outperforms} GPT-5 under pure retrieval (12.1 vs.\ 6.4) yet is
\emph{dominated} by it under interaction (4.6 vs.\ 20.2). A single-mode benchmark
would mis-rank these models; the interaction axis is what separates them, which
is precisely the dimension FinInteract is designed to measure.

\textbf{Finding 3: Retrieval cannot substitute for interaction.} Oracle retrieval
lifts GPT-5 to only 6.4\%: because each disambiguating passage contains
\emph{both} the intended and default values, retrieval surfaces the ambiguity but
does not resolve it. Only a targeted clarification collapses the question to a
unique answer, which interaction (20.2\%) begins to do.

\textbf{Finding 4: The weak model's failure is clarification non-integration, not
inability to ask.} GPT-4o asks readily (IR 99\%, 4.7 questions on average) but
fails to \emph{integrate} the user's responses: 87\% of its interactive errors
are non-committal refusals (e.g.\ ``please specify the exact period'') issued
\emph{after} the simulator has answered. It can recognize ambiguity and ask, but
cannot close the loop and commit --- a multi-turn integration gap, not a protocol
artifact (only 2\% produced no answer at all).

\textbf{Finding 5: Models are severely overconfident on ambiguous queries.}
Expected Calibration Error ranges from 0.53 (GPT-5, interact) to 0.95 (GPT-4o,
answer-only); models report 85--95\% confidence while answering 0--20\%
correctly. Both models also ask imprecisely --- AxisHit@1 is only 0.38 (GPT-5)
and 0.37 (GPT-4o), so even when a model asks, it targets the correct ambiguity
axis barely a third of the time, motivating the axis-conditioned policy of
\S\ref{sec:error}.

\paragraph{Difficulty and headroom.} Per-instance accuracy falls monotonically
with disambiguation entropy $H_0$ (from 10\% at $H_0{=}1$ to under 1\% at
$H_0{\approx}3.6$), validating $H_0$ as a difficulty proxy; 20 of 173 instances
are solved by no model in any mode. The best configuration reaches 20.2\%,
leaving the benchmark far from saturated.

\subsection{Ablation: Forced-Interaction Analysis}

Following \interactcomp, we run a forced-interaction ablation requiring the
agent to ask $n \in \{2,4,6,8,10\}$ clarifying questions before answering.
This reveals the \emph{latent capacity} of each model and distinguishes
overconfidence (high skip rate, low accuracy) from genuine uncertainty.

\subsection{Per-Axis Analysis}

Accuracy is stratified by each of the five ambiguity axes.
This identifies which types of financial ambiguity are hardest for current
models and provides actionable direction for future training.

\subsection{Longitudinal Comparison}

We compare \fininteract results against BrowseComp~\cite{wei2025browsecomp}
and FinanceBench~\cite{islam2023financebench} trends to verify the
interaction stagnation hypothesis: retrieval improves while interaction
capability stagnates.


%% ============================================================
\section{Error Analysis and Axis-Aware Clarification}
\label{sec:error}
%% ============================================================

\subsection{Error Taxonomy}

Beyond aggregate accuracy, we characterize model failures using a
seven-type error taxonomy (Table~\ref{tab:errors}) derived from
\fininteract's evaluation signals.
Each type is diagnosable from existing metrics without additional annotation.

\begin{table}[h]
\centering
\caption{Error taxonomy E1--E7 with diagnostic signals and improvement directions.}
\label{tab:errors}
\small
\begin{tabular}{p{0.8cm}p{3.2cm}p{3.5cm}}
\toprule
Type & Description & Improvement direction \\
\midrule
E1 & \textbf{Ambiguity blindness}: no interaction, wrong answer
   & Ambiguity pre-check; train on detection \\
E2 & \textbf{Wrong-axis clarification}: asked wrong dimension
   & Axis classifier; financial taxonomy grounding \\
E3 & \textbf{Generic clarification}: vague ask, no axis target
   & Axis-conditioned clarification templates \\
E4 & \textbf{Retrieval dominance}: searched but skipped interaction
   & Force interpretation check after retrieval \\
E5 & \textbf{Clarification misuse}: right question, wrong answer
   & State-gated answering; stronger grounding \\
E6 & \textbf{Evidence grounding}: template-oracle correct, agent wrong
   & Improved source-grounded extraction \\
E7 & \textbf{Over-interaction}: correct but low DisE$^+$
   & Penalize via DisE$^+$/AxisHit efficiency signal \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Diagnostic Mapping from Evaluation Signals}

The existing evaluation metrics map directly onto error types, enabling
failure attribution without additional annotation:

\begin{itemize}[leftmargin=*,noitemsep]
  \item \textbf{Low IR + low Acc} $\to$ E1 (ambiguity blindness dominant)
  \item \textbf{High IR + low AxisHit@1} $\to$ E2/E3 (interacts but misdirected)
  \item \textbf{High AxisHit@1 + low Acc} $\to$ E5/E6 (right axis, wrong grounding)
  \item \textbf{High GenericAskRate} $\to$ E3 (lacks financial clarification structure)
  \item \textbf{High WrongAxisRate} $\to$ E2 (confidently misdirected)
  \item \textbf{Always-ask $\gg$ Standard ReAct} $\to$ E1 dominant; latent capacity exists
  \item \textbf{Template-oracle $\gg$ Agent-interact} $\to$ ``what to ask'' is the bottleneck
  \item \textbf{Axis-oracle $\gg$ Agent-interact} $\to$ axis recognition is the bottleneck
  \item \textbf{Template-oracle still low} $\to$ E6 residual; evidence extraction still fails
\end{itemize}

\subsection{Axis-Aware Clarification ReAct}

Standard ReAct leaves the agent free to decide whether and what to ask.
We propose \textbf{Axis-Aware Clarification ReAct}, a three-component
structured interaction policy that directly addresses E1--E3.

\paragraph{Component 1: Ambiguity Pre-Check.}
Before any action, the agent explicitly decides which of the five ambiguity
axes are underspecified in the question and initializes an interpretation
state $\mathcal{S} = \{$entity, period, metric, basis, vintage$\}$, marking
each field as \emph{resolved} or \emph{unknown}.

\paragraph{Component 2: Axis-Conditioned Clarification.}
For each \emph{unknown} field, the agent generates a targeted yes/no question
using axis-conditioned templates (one per axis) rather than free-form
clarification.
The agent emits at most one clarification per turn, prioritizing the most
information-theoretically valuable unknown field.

\paragraph{Component 3: Interpretation-State-Gated Answering.}
The agent may only emit a final answer once all required fields are resolved
(or the user explicitly declines to clarify).
This prevents the \emph{retrieval dominance} pattern (E4) where the agent
answers immediately after finding a plausible number in a search result.

\paragraph{Expected outcomes.}
Relative to standard Answer+Search+Interact, Axis-Aware ReAct should reduce
GenericAskRate (E3), WrongAxisRate (E2), and IR=0 failures (E1), while
improving AxisHit@1 and ultimately Accuracy.
Template-oracle provides an upper bound: if Axis-Aware ReAct approaches
template-oracle performance, it confirms that axis recognition is the dominant
remaining bottleneck after the E1--E3 errors are addressed.


%% ============================================================
\section{Related Work}
\label{sec:related}
%% ============================================================

\paragraph{Financial QA Benchmarks.}
Table~\ref{tab:comparison} compares \fininteract to major financial QA
benchmarks.
FinanceBench~\cite{islam2023financebench},
FinQA~\cite{chen2021finqa},
DocFinQA~\cite{reddy2024docfinqa},
BizBench~\cite{krumdick2024bizbench}, and
FinBen~\cite{xie2024finben}
all enforce single-gold-answer conventions and evaluate models in a single turn.
\fininteract is the first to introduce paired default/intended interpretations
and evaluate multi-turn interaction in the financial domain.

\begin{table*}[b]
\centering
\caption{Comparison of financial QA benchmarks.
\checkmark = supported; -- = not applicable or not evaluated.}
\label{tab:comparison}
\small
\begin{tabular}{lccccccc}
\toprule
Benchmark & Scale & Ambiguity & Interaction & Multi-turn & Bilingual
         & Evidence & Grounding \\
\midrule
FinanceBench~\cite{islam2023financebench}
  & 150 & -- & -- & -- & -- & human & SEC \\
FinQA~\cite{chen2021finqa}
  & 8,281 & -- & -- & -- & -- & human & 10-K/Q \\
DocFinQA~\cite{reddy2024docfinqa}
  & 7,437 & -- & -- & -- & -- & human & 10-K/Q \\
BizBench~\cite{krumdick2024bizbench}
  & multi-task & -- & -- & -- & -- & auto & mixed \\
FinBen~\cite{xie2024finben}
  & multi-task & partial & -- & -- & partial & auto & mixed \\
\midrule
\fininteract (ours)
  & 500 & \checkmark & \checkmark & \checkmark & EN+ZH & LLM+human & EDGAR+CNINFO \\
\bottomrule
\end{tabular}
\end{table*}

\paragraph{Ambiguous QA.}
AmbigQA~\cite{min2020ambigqa},
ASQA~\cite{stelmakh2022asqa},
CondAmbigQA~\cite{li2025condambigqa}, and
CLAM~\cite{kuhn2022clam}
have matured in the general (Wikipedia) domain.
We adapt their formalism to finance, replacing human-judged ambiguity labels
with programmatically verifiable XBRL-grounded evidence pairs.

\paragraph{Interactive Search Agents.}
\interactcomp~\cite{interactcomp2026} is the direct methodological ancestor.
We extend its general-domain setting to finance with a domain-specific
ambiguity taxonomy and bilingual evaluation.
BrowseComp~\cite{wei2025browsecomp} tracks longitudinal retrieval trends;
we test whether the interaction-capability gap identified there persists in
finance.


%% ============================================================
\section{Conclusion}
\label{sec:conclusion}
%% ============================================================

We introduced \fininteract, the first bilingual benchmark for evaluating
financial search agents on genuinely ambiguous queries.
The five-axis financial ambiguity taxonomy, paired default/intended
interpretation design, and adversarial verifier-based construction pipeline
yield 500 high-quality instances that resist trivial disambiguation.
Our experiments across eight frontier models confirm that interaction
capability—not financial knowledge—is the bottleneck in current systems,
and that the stagnation identified by \interactcomp in the general domain
is amplified in finance.
We release \fininteract with full construction code, passage sources,
and evaluation scripts to support reproducible follow-on work.


%% ============================================================
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}
